### 2019-10-07 tess.tex
Final version

TTV studies must continue the work of space missions from the ground. It is unlike, only with space-based observations, to cover the interaction timescales to dynamically prove the TTV signal and the existence of possible companions, as in shown in Figure \ref{koinet}. In this figure, new data from ground-based telescopes were fundamental to confirm the TTV signal, and therefore to perform a proper characterization of the system.

@@

The new space telescope Transiting Exoplanet Survey Satellite (TESS: \cite{Ricker2014}) is providing its first discoveries. The TESS observation plan divided into 13 sectors will challenge the confirmation process and characterization of possible exoplanets. During the $\sim27$ days of each sector, many candidates will not be observed enough time to confirm and characterize them. However, this special feature will provide a huge amount of new transit events of already known transiting exoplanets, especially for those with short-period. Combining the new TESS data with ground-based follow- up observations, many possible TTVs could be confirmed or ruled out. Moreover, \cite{Ballard2018} predicted that around 5\% of the planets discovered by TESS will exhibit measurable TTV signals, similar to Kepler’s results.

Between Sector 1 and Sector 13, TESS observed around 15\% (see Figure \ref{match_tess}) of the stars with confirmed exoplanets but not all of them will produce a transit event within the 27 days sector's duration (see Figure \ref{chap:tess}). Anyway, several complete light curves could be obtained and analyzed. TESS data are reduced by the Science Processing Operations Center (SPOC) and after being processed, they are archived in the Mikulski Archive for Space Telescopes (MAST) catalog, where can be downloaded directly by anyone.

@@

In the following subsections, I present a re-analysis of the TTV curves of the hot Jupiters: WASP-18b, WASP-19b, and WASP-77Ab including new transit times from TESS: 45, 28, and 15 for WASP-18b, WASP-19b and WASP-77Ab respectively. To compute the transit time of each light curve, I used EXOFASTv2 to only fit the transit time and the flux's baseline. The remaining planetary parameters were fixed with the values from Tables \ref{wasp18}, \ref{tab:wasp19}, \ref{tab:wasp77}.

@@

The new best linear fit for the period considering a total of 63 transit times is:

\begin{equation}
P = 0.941452232 \pm 2.7\times10^{-8} ~\rm days
\end{equation}

This value differs from the previously reported in Eq.~\ref{eq1_w18} by $0.016$ seconds, and the error was reduced to a third.

Figure \ref{wasp18_ttv} shows the extension of the TTV curve of WASP-18b considering the new data from TESS. Including these new transit times, the scatter from the linear ephemeris was reduced to 47 seconds. A possible second-order fit of a hypothetical orbital decay has a reduced chi-squared of $\chi^{2}_{red}=0.35$, while the linear fit - which corresponds to the linear ephemeris - has $\chi^{2}_{red}=0.36$. Furthermore, as there is no strong evidence for orbital decay in this system as proposed by \cite{Wilkins2017}, the simpler model should be chosen: the linear model.

@@

\begin{figure}[h]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP18_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-18b. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 63 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$. The RMS scatter from the linear ephemeris is 47 seconds.}
\label{wasp18_ttv}
\end{figure}

@@

WASP-19 was observed during Sector 9 between February 28th and March 26th, 2019. During this time, 29 complete transit event was provided and then considered to compute the new TTVs of WASP-19b shown in Figure \ref{wasp19_ttv}. The new best fit for the orbital period P, considering a total of 88 transit times is:

\begin{equation}
P = 0.788838940 \pm 3.0\times10^{-8} ~\rm days
\end{equation}

The previously reported value in Eq.~\ref{eq1_wasp19} is $0.007$ seconds larger.

As in the case of WASP-18b, the TTV scatter of WASP-19b was reduced when including new transit times data from TESS to: $TTV_{RMS}=65~\rm seconds$. Again, a second order fit is not better that the linear one: $\chi^{2}_{red,lineal}= 0.65$ and $\chi^{2}_{red,cuad}= 0.64$.

\begin{figure}[ht]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP19_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-19b. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 87 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$. The RMS scatter from the linear ephemeris is 65 seconds.}
\label{wasp19_ttv}
\end{figure}

@@

The third target presented in Cortés-Zuleta et al. 2019 (submitted), was observed in Sector 4 between October 18th and November 15th, 2018.  During this time, the planet transited its host star 17 times, but only 15 were complete light curves. From the previous study presented in this thesis, WASP-77Ab was the most interesting target, since its variation from the linear ephemeris was around  $121\,\rm seconds$.  After including the new transit times, the best linear fit for the orbital period is:

\begin{equation}
P = 1.36002866 \pm 1.7\times10^{-7} ~\rm days
\end{equation}

In Eq.~\ref{eq1_w77}, the value for the orbital period P is $0.01\,\rm seconds$ lower, and the scatter from the linear ephemeris was reduced to 86 seconds. The linear fit has a reduced chi-squared of $\chi^2_{red}=1.03$, while a second-degree polynomial has $\chi^2_{red}=0.72$. Anyway, the second-order fit is highly dominated by the outlier transit time of epoch $229$ (see Figure \ref{wasp77_ttv}). After removing it, the reduced chi-squared is  $\chi^2_{red}=0.37$. In all cases, the best fit corresponds to the linear ephemeris.

@@

\begin{figure}[ht]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP77_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-77Ab. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 26 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$. The RMS scatter from the linear ephemeris is 86 seconds.}
\label{wasp77_ttv}
\end{figure}

>>>>

TTV studies must continue the work of space missions from the ground. Ground-based photometric follow-up observations are important and contribute with significant scientific measurements and complement space-based observations to detect astrophysical interesting signals like TTVs. In Figure \ref{koinet}, new data from ground-based telescopes were fundamental to confirm the TTV signal, and therefore to perform a proper characterization of the system.

@@

The new space telescope Transiting Exoplanet Survey Satellite (TESS: \cite{Ricker2014}) is delivering its first discoveries. The TESS observation plan divided into 13 sectors will challenge the confirmation process and characterization of possible exoplanets. During the $\sim27$ days of each sector, many candidates will not be observed for a long enough time to confirm and characterize them. However, this special feature will provide a huge amount of new transit events of already known transiting exoplanets, especially for those with short-period. Combining the new TESS data with ground-based follow- up observations, many possible TTVs could be confirmed or ruled out. Moreover, \cite{Ballard2018} predicted that around 5\% of the planets discovered by TESS will exhibit measurable TTV signals, similar to Kepler’s results.

Between Sector 1 and Sector 13 -- sectors dedicated to the southern hemisphere -- TESS observed around 15\% (see Figure \ref{match_tess}) of the stars with confirmed exoplanets but not all of them will produce a transit event within the 27 days sector's duration (see Figure \ref{chap:tess}). Anyway, several complete light curves could be obtained and analyzed. TESS data are reduced by the Science Processing Operations Center (SPOC: \cite{Jenkins2016}) and after being processed, they are archived in the Mikulski Archive for Space Telescopes (MAST\footnote{http://archive.stsci.edu/}) catalog, where can be downloaded directly by anyone.

@@

In the following subsections, I present a re-analysis of the TTV curves of the hot Jupiters: WASP-18b, WASP-19b, and WASP-77Ab including new transit times from TESS: 45, 28, and 15 for WASP-18b, WASP-19b and WASP-77Ab respectively. To compute the transit time of each light curve, I used EXOFASTv2\footnote{https://github.com/jdeast/EXOFASTv2} \citep{Eastman2013} to only fit the transit time and the flux's baseline. The remaining planetary parameters were fixed with the values from Tables \ref{wasp18}, \ref{tab:wasp19}, \ref{tab:wasp77}.

@@

The new linear ephemeris equation considering a total of 63 transit mid-times is:

\begin{equation}
T_{\rm C}(E)=2456926.27460\pm(94)+E \cdot 0.941452232 \pm(27)
\end{equation}

The best fit for the period $P= 0.941452232 \pm 2.7\times10^{-8} ~\rm days$ differs from the previously reported in Eq.~\ref{eq1_w18} by $0.016$ seconds, and the error was reduced to a third.

Figure \ref{wasp18_ttv} shows the extension of the TTV curve of WASP-18b considering the new data from TESS. Including these new transit times, the scatter from the linear ephemeris was reduced to 47 seconds. A possible second-order fit of a hypothetical orbital decay has a reduced chi-squared of $\chi^{2}_{red,quad}=0.35$, while the linear fit - which corresponds to the linear ephemeris - has $\chi^{2}_{red,lin}=0.36$. Furthermore, as there is no strong evidence for orbital decay in this system as proposed by \cite{Wilkins2017}, the simpler model should be chosen: the linear model.

@@

\begin{figure}[h]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP18_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-18b. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 63 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$, which follows a imperceptible quadratic trend. The RMS scatter from the linear ephemeris is 47 seconds.}
\label{wasp18_ttv}
\end{figure}

@@

WASP-19 was observed during Sector 9 between February 28th and March 26th, 2019. During this time, 29 complete transit event was provided and then considered to compute the new TTVs of WASP-19b shown in Figure \ref{wasp19_ttv}. The new proposed linear ephemeris equation considering a total of 88 transit times is:

\begin{equation}
T_{\rm C}(E)=2456402.7128\pm(16)+E \cdot 0.788838940 \pm(30)
\end{equation}

The new best fit value for the period $P=0.788838940 \pm 3.0\times10^{-8} ~\rm days$ is $0.007$ seconds larger than the previously reported value in Eq.~\ref{eq1_wasp19}.

As in the case of WASP-18b, the TTV scatter of WASP-19b was reduced when including new transit times data from TESS to: $TTV_{RMS}=65~\rm seconds$. Again, a second order fit is not better that the linear one: $\chi^{2}_{red,lin}= 0.65$ and $\chi^{2}_{red,quad}= 0.64$.

\begin{figure}[ht]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP19_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-19b. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 87 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$, which follows a imperceptible quadratic trend.. The RMS scatter from the linear ephemeris is 65 seconds.}
\label{wasp19_ttv}
\end{figure}

@@

The third target presented in Cortés-Zuleta et al. 2019 (submitted), was observed in Sector 4 between October 18th and November 15th, 2018.  During this time, the planet transited its host star 17 times, but only 15 were complete light curves. From the previous study presented in this thesis, WASP-77Ab was the most interesting target, since its variation from the linear ephemeris was around  $121\,\rm seconds$.  After including the new transit times, the new proposed linear ephemeris equation is:

\begin{equation}
T_{\rm C}(E)=2457420.88439\pm(85)+E \cdot 1.36002866 \pm(17)
\end{equation}

With a best fit period of $P=1.36002866 \pm 1.7\times10^{-7} ~\rm days$. In Eq.~\ref{eq1_w77}, the value for the orbital period P is $0.01\,\rm seconds$ lower, and the scatter from the linear ephemeris was reduced to 86 seconds. The linear fit has a reduced chi-squared of $\chi^2_{red,lin}=1.03$, while a second-degree polynomial has $\chi^2_{red,quad}=0.72$. Anyway, the second-order fit is highly dominated by the outlier transit time of epoch $229$ (see Figure \ref{wasp77_ttv}). After removing it, the reduced chi-squared is  $\chi^2_{red,quad}=0.37$. In all cases, the best fit corresponds to the linear ephemeris.

@@

\begin{figure}[ht]
%\vspace{0.1cm}\hspace{0.2cm}
\includegraphics[width=1.0\textwidth]{imagenes/WASP77_TTV.pdf}
\caption{Observed minus calculated mid-transit times (TTV) for WASP-77Ab. The dashed black line corresponds to the proposed linear ephemeris, i.e. zero deviation from the predicted transit time  computed from our refined orbital period. For that, we considered 26 transit times from the TraMoS project (in light blue), published works and TESS (both in black). The grey area corresponds to the error propagation at $1\sigma$, which follows a imperceptible quadratic trend. The RMS scatter from the linear ephemeris is 86 seconds.}
\label{wasp77_ttv}
\end{figure}

### 2019-10-07 tramos.tex
Final version

The Transit Monitoring in the South (TraMoS) project started in 2008 aiming to perform a photometric follow-up of transiting exoplanets using telescopes located in Chile.  Through photometric monitoring of transiting events, important information of the planetary system can be obtained, such as the planetary radius, inclination of the orbit, precise planetary mass, among others. Moreover, this kind of analysis allows the detection of variability in the transit's parameters (TPV), if several epochs of the transit event are included. The TraMoS project aims, on one hand, to refine the physical and orbital parameters of selected exoplanets systems through photometric follow-up and, on the other, to search variations in those parameters that could suggest the presence of additional bodies in the system.

To date, almost 400 transit events have been observed of 144 transiting exoplanets (see Figure \ref{tramos}). The bulk of the targets in the TraMoS project are hot Jupiters due to its short orbital period and large transit depth. Several facilities were used to perform the photometric follow-up such as the VLT of European Southern Observatory, the SMARTS 0.9 m and 1 m at Cerro Tololo Inter-American Observatory, the Danish 1.54 m at La Silla Observatory, SWOPE and Du Pont telescopes at Las Campanas Observatory and SOAR telescope at Cerro Pachón Observatory, among others. The diameter size of the used telescopes ranges from 0.6 meters to 8 meters. However, in the last stage of the TraMoS project, we built the team expertise using 1 meter-class telescope, for example in the Danish, SWOPE and SMARTS.

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{imagenes/tramos.png}
\caption{Histogram of the 30 transiting exoplanets with the largest amount of observations within the TraMoS project. The exoplanets WASP-4b and WASP5-b were presented in \cite{Hoyer2012} and \cite{Hoyer2013}, respectively. The results from the analysis of WASP-18b, WASP-19b and WASP77-Ab are in Cortés-Zuleta et al. 2019 (submitted, see Chapter \ref{chap:paper} for further details). In none of those systems a significant TTV signal was detected.}
\label{tramos}
\end{figure}

The first studies within the TraMoS project were conducted by Sergio Hoyer, former PhD student of Professor Patricio Rojo. In summary, a possible orbital decay of WASP-43b \citep{Hoyer2016b} and OGLE-TR-113b \citep{Hoyer2016} was ruled out, and no significant TTV signal was detected in WASP-4b \citep{Hoyer2012} and WASP-5b \citep{Hoyer2013}.

In the following section \ref{lightcurve}, I describe the light curve produced from a transiting exoplanet --the principal target of the TraMoS project --  what information about the planet can be obtained from them and how these light curves can give us information about the architecture of this kind of systems. Then, in section \ref{ttv}, I detail on a specific type of phenomena of transiting exoplanets: the variation of the transit time of its light curve (TTV).

@@

In Figure~\ref{transit} is described the architecture of a transiting exoplanet. Thanks to its geometry, the transit light curve and the parameters of the planetary system can be described with a few simple relations \cite{Seager2003,Winn2010}.  When the planet passes in front of its star produced a dimming in its flux called primary transit. The maximum drop during the transit event defines the transit depth $\Delta{F}$:

@@

Where $R_{p}$ and $R_{*}$ are the planet and star radius, respectively. The numbers 1, 2, 3 and 4 correspond to the contact times: $t_{1}$, $t_{2}$, $t_{3}$ and $t_{4}$. The total duration of the transit is $t_{T} = t_{4}-t_{1}$ and the full duration is $t_{F}=t_{3}-t_{2}$. The total duration of the transit can be also described considering planetary parameters such as the period $P$, semi-major axis $a$ and orbital inclination $i$:

@@

Where $\omega$ is the periastron longitude. In the case of circular orbits, the eccentricity is zero, thus Eq.~\ref{impact_param} is simplified to:

@@

\begin{figure}[ht]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/transit_architecture.pdf}
\caption{Architecture of a transiting exoplanet and the observed light curve. Based on the light curve itself it is possible to measure the transit depth $\Delta F$ and the impact parameter $b$. From them, the planet-to-star radius ratio $R_{p}/R_{*}$ and the orbital inclination $i$ can be derived. Figure from The Handbook of Exoplanets by M. Perryman.}
\label{transit}
\end{figure}

Depending on the orbital and physical parameters of the exoplanet, the shape of the observed light curve will vary. For example, Figure \ref{transit_examples} shows how the light curve reflects the configuration of the system. Hot Jupiters are the kind of planet easier to detect since their planet-to-star radius ratio is larger, but a smaller planet orbiting a dwarf star could produce a similar ratio of the radius. On the other hand, exoplanets with larger orbital inclinations will produce shorter drops in the flux, in comparison with planets in the same orbital-plane with the host star.

@@

When modeling the light curve's data, not only the physical parameters of the planet need to be taken into account, but the stars' too. The Limb Darkening effect of the star will change the shape of the observed light curve, as what we are observing are photons coming from the star itself, especially when carrying-out observations with different filters. This effect produces that the stars appear to be brighter in the center of it and darker closer to the edge. The most general analytic formulae to model a light curve produced by a transiting exoplanet includes the Limb Darkening coefficients \citep{Mandel2002}, which can be quadratic or non-quadratic.

@@

Since the discovery of the first exoplanet, one of the principal questions that arose was: are they alone in their systems? The idea of multi-planetary systems is a consequence of our knowledge about the planetary system where we live in, and until today the Solar System has a larger number of orbiting planets. Besides the Solar System, Kepler-90 has the same number of confirmed exoplanets: eight. Statistically, around a 20\% of the confirmed exoplanets live in a multi-planetary system\footnote{Information obtained from the NASA Exoplanet Archive Database \url{exoplanetarchive.ipac.caltech.edu}}, but this number is theoretically larger. \cite{Fressin2013} computed the occurrence of exoplanets through numerical simulation with Kepler data, obtaining that around $52\%$ of the stars should have at least one planet within an 85 days orbit.

The Transit Timing Variations (TTV) technique was first proposed as a way to detect Earth-mass planets in multi-planetary systems due to gravitational interactions with a transiting exoplanet \citep{Holman2005,Agol2005}. However, TTVs can be also used to measure tidal interactions between the planet and the star, such as orbital decays for instance, or to detect exomoons \citep{Kipping2009a,Kipping2009b}, among others phenomena in the star itself.

For a given light curve of a transiting exoplanet (see Figure \ref{transit}), the transit time $T_{c}$ (or central time of the transit) is computed as:

\begin{equation}
T_{c} = \frac{t_T}{2}
\end{equation}

A single transiting planet will stay in a Keplerian orbit around its host star, producing a transit time strictly periodic. Thus, the transit time $T_c$ follows a linear function of the orbital period $P$:

@@

Where $T_0$ is a reference transit time and $E$ is the number of epochs since $T_0$.  The presence of a second planet produces transits no exactly periodic, hence Eq.~\ref{linear_ephemeris} will not be valid. As the transit times will be no longer a linear function,  the time between transits varies.  The variation in time produced by the perturber body depends on the mass and geometry of their orbits. The magnitude of this variation between successive transits of the transiting planet \citep{Holman2005}, assuming that the perturber body is in a inner orbit, $a_2 < a_1$, and follows a parabolic orbit with a pariastron distance of $a_2(1-e_2)$, is:

\begin{equation}
\Delta t = \frac{45\pi}{16} \frac{M_2}{M_*} P_{1} \alpha^3_{e} (1-	\sqrt{2}\alpha^{3/2})^{-2}
\label{delta_t}
\end{equation}

Where $\alpha_e = a_1/(a_2(1-e_2))$ is the ratio of the semi-major axis of the planets, considering a nonzero eccentricity for the perturbing planet ($e_2 \neq 0$).The orbital period of transiting exoplanet is $P_1$, while the perturbing planet has a mass $M_2$, and orbital period $P_2$. When $e_2>0.3$, the relation \ref{delta_t} has the best performance with numerical simulations.

@@

\cite{Steffen2012b} analyzed Kepler data to constraint the occurrence rate of companions in hot Jupiter system. In a sample of 63 hot Jupiter candidates, none of them show evidence of possible companions in their system. When comparing with other types of exoplanets, 30\% of hot Neptune candidates may have companions and 10\% of warm Jupiter candidates. This could suggest that hot Jupiter may have different formation processes on which smaller companions are not allowed in their system: probably they are ejected through planet-planet scatter during migration. TTV studies could provide important clues about formation processes and dynamical evolution, and their mass and period dependence.

>>>>

The Transit Monitoring in the South (TraMoS) project started in 2008 aiming to perform photometric follow-up of transiting exoplanets using telescopes located in Chile.  Through photometric monitoring of transiting events, important information of the planetary system can be obtained, such as the planetary radius, inclination of the orbit, and, in combination with radial velocity data, the planetary mass, among others. Moreover, this kind of analysis allows the detection of variability in the transit's parameters (TPV), if several epochs of the transit event are obtained. The TraMoS project aims, on one hand, to refine the physical and orbital parameters of selected exoplanets systems through photometric follow-up and, on the other, to search for variations in those parameters that could suggest the presence of additional bodies in the system.

To date, almost 400 transit events have been observed of 144 transiting exoplanets (see Figure \ref{tramos}). The bulk of the targets in the TraMoS project are hot Jupiters due to its short orbital period and large transit depth. Several facilities were used to perform the photometric follow-up such as the VLT of European Southern Observatory, the SMARTS 0.9 m and 1 m at Cerro Tololo Inter-American Observatory, the Danish 1.54 m at La Silla Observatory, SWOPE and Du Pont telescopes at Las Campanas Observatory and SOAR telescope at Cerro Pachón Observatory, among others. The diameter size of the used telescopes ranges from 0.6 meters to 8 meters. However, in the later stage of the TraMoS project, we built the team expertise using 1 meter-class telescope, for example in the Danish, SWOPE and SMARTS.

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{imagenes/tramos.png}
\caption{Histogram of the 30 transiting exoplanets with the largest amount of observations within the TraMoS project. The exoplanets WASP-4b and WASP5-b were presented in \cite{Hoyer2012} and \cite{Hoyer2013}, respectively. The results from the analysis of WASP-18b, WASP-19b and WASP77-Ab are in Cortés-Zuleta et al. 2019 (submitted, see Chapter \ref{chap:paper} for further details). In none of those systems a significant TPV signal was detected.}
\label{tramos}
\end{figure}

The first studies within the TraMoS project were conducted for Segio Hoyer's PhD dissertation (advisor: Patricio Rojo). Current results are summarized as: a possible orbital decay of WASP-43b \citep{Hoyer2016b} and OGLE-TR-113b \citep{Hoyer2016} was ruled out, and no significant TTV signal was detected in WASP-4b \citep{Hoyer2012} and WASP-5b \citep{Hoyer2013}.

In the following section \ref{lightcurve}, I describe the light curve produced from a transiting exoplanet --the principal target of the TraMoS project --  what information about the planet can be obtained from them and how these light curves can give us information about the architecture of this kind of systems. Then, in section \ref{ttv}, I detail on a specific type of phenomena of transiting exoplanets: the variation of the transit mid-time of its light curve (TTV).

@@

In Figure~\ref{transit} is shown the geometry of a transiting exoplanet. From geometric considerations the transit light curve and the parameters of the planetary system can be described with a few simple relations \cite{Seager2003,Winn2010}.  When the planet passes in front of its star produced a dimming in its flux called primary transit. The maximum drop during the transit event defines the transit depth $\Delta{F}$:

@@

Where $R_{p}$ and $R_{*}$ are the planet and star radius, respectively. The numbers 1, 2, 3 and 4 correspond to the contact times: $t_{1}$, $t_{2}$, $t_{3}$ and $t_{4}$ (see. Fig. \ref{transit}). The total duration of the transit is $t_{T} = t_{4}-t_{1}$ and the full duration is $t_{F}=t_{3}-t_{2}$. The total duration of the transit can be also described considering planetary parameters such as the period $P$, semi-major axis $a$ and orbital inclination $i$:

@@

Where $\omega$ is the argument of pericenter. In the case of circular orbits, the eccentricity $e$ is zero, thus Eq.~\ref{impact_param} is simplified to:

@@

\begin{figure}[ht]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/transit_architecture.pdf}
\caption{Geometry of a transiting exoplanet and the observed light curve. Based on the light curve itself it is possible to measure the transit depth $\Delta F$ and the orbit's inclination $i$. The numbers in circle indicate the four contact times: $t_{1}$, $t_{2}$, $t_{3}$ and $t_{4}$.  Figure from The Handbook of Exoplanets by M. Perryman.}
\label{transit}
\end{figure}

Depending on the orbital and physical parameters of the exoplanet, the shape of the observed light curve will vary. For example, Figure \ref{transit_examples} shows how the light curve reflects the configuration of the system. Hot Jupiters are the kind of planet that is relatively easier to detect since their planet-to-star radius ratio is larger, but a smaller planet orbiting a dwarf star could produce a similar ratio of the radius. On the other hand, exoplanets with smaller orbital inclinations will produce shallower drops in the flux, in comparison with planets in the same orbital-plane with the host star.

@@

When modeling the light curve's data, not only the physical parameters of the planet need to be taken into account, but the stars' too. The Limb Darkening effect of the star will change the shape of the observed light curve (during the time period $t3 - t2$ in Fig.~\ref{transit}), as what we are observing are photons coming from the star itself, especially when carrying-out observations with different filters. The effect of stellar limb-darkening describes the intensity variation across the disk of the star: the star appears brighter at the center and darker closer to the edge (stellar limb). The most general analytic formulae to model a light curve produced by a transiting exoplanet includes the Limb Darkening coefficients \citep{Mandel2002}, which can be quadratic or non-quadratic.

@@

Since the discovery of the first exoplanet, one of the principal questions that arose was: are they alone in their systems? The idea of multi-planetary systems is a consequence of our knowledge about the planetary system where we live in, and until today the Solar System has a larger number of orbiting planets. Besides the Solar System, Kepler-90 has the same number of confirmed exoplanets: eight. Statistically, around 20\% of the confirmed exoplanets are part of a multi-planetary system\footnote{Information obtained from the NASA Exoplanet Archive Database \url{https://exoplanetarchive.ipac.caltech.edu}}, but this number is theoretically larger. \cite{Fressin2013} computed the occurrence of exoplanets through numerical simulation with Kepler data, obtaining that around $52\%$ of the stars should have at least one planet within an 85 days orbit.

The Transit Timing Variations (TTV) technique was first proposed as a way to detect Earth-mass planets in multi-planetary systems due to gravitational interactions with a transiting exoplanet \citep{Holman2005,Agol2005}. However, as mention in Chapter \ref{chap:intro}, TTVs can be also used to measure tidal interactions between the planet and the star, such as orbital decays for instance, or to detect exomoons \citep{Kipping2009a,Kipping2009b}, among others phenomena in the star itself.

For a given light curve of a transiting exoplanet (see Figure \ref{transit}), the transit mid-time $T_{c}$ (or central time of the transit) is computed as:

\begin{equation}
T_{c} = \frac{t_{4}+t_{1}}{2}
\end{equation}

A single (non-perturbed) transiting planet will stay in a Keplerian orbit around its host star, producing a transit time strictly periodic. Thus, the transit time $T_c$ follows a linear function in the transit epoch $E$:

@@

Where $T_0$ is a reference transit time, $E$ is the number of epochs since $T_0$, and $P$ is the orbital period of the planet. The presence of a second (perturbing) planet produces transits no exactly periodic, hence Eq.~\ref{linear_ephemeris} will not be valid. As the transit times will be no longer a linear function,  the time between transits varies.  The variation in time produced by the perturber body depends on the mass and geometry of their orbits. The magnitude of this variation between successive transits of the transiting planet \citep{Holman2005}, assuming that the perturber body is in a inner orbit, $a_2 < a_1$, and follows a eccentric orbit with a periastron distance of $a_2(1-e_2)$, is:

\begin{equation}
\Delta t = \frac{45\pi}{16} \frac{M_2}{M_*} P_{1} \alpha^3_{e} (1-	\sqrt{2}\alpha_{e}^{3/2})^{-2}
\label{delta_t}
\end{equation}

Where $\alpha_e = a_1/(a_2(1-e_2))$ is the ratio of the semi-major axis of the planets, considering a nonzero eccentricity for the perturbing planet ($e_2 \neq 0$).The orbital period of transiting exoplanet is $P_1$, while the perturbing planet has a mass $M_2$, and orbital period $P_2$. The Eq. \ref{delta_t} obtains the best results, compared with numerical simulations, when $e_2>0.3$.

%When $e_2>0.3$, the relation \ref{delta_t} has the best performance with numerical simulations.

@@

\cite{Steffen2012b} analyzed Kepler data to constrain the occurrence rate of companions in hot Jupiter system. In a sample of 63 hot Jupiter candidates, none of them show evidence of possible companions in their system. When comparing with other types of exoplanets, 30\% of hot Neptune candidates may have companions and 10\% of warm Jupiter candidates. This could suggest that hot Jupiter may have different formation processes on which smaller companions are not allowed in their system: probably they are ejected through planet-planet scatter during migration. TTV studies could provide important clues about formation processes and dynamical evolution, and their mass and period dependency.
